---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

- Each breeding adult will produce, on average, 600 eggs
- 5% of eggs will survive to become fry
- 20% of fry will survive to become juveniles
- 16% of juveniles will survive to become adults
- Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. 

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600,
                        0.05, 0, 0, 0,
                        0, 0.20, 0, 0,
                        0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by: 
```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
```


Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step. 

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ------ comment line-by-line ---- #

# The line below creates the time vector by year, representing the lifecycles
time_yr <- seq(from = 0, to = 8, by = 1)

# The line below creates empty matrix with ncols 4 and nrows 8
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# The line below fills in empty, first row with initial population
proj_year[1, ] <- fish_year0

# The for loop below calculates dot product of 'survival' (leslie) and proj_year, calculating populations of each lifestage and filling in empty row

for (i in 2:length(time_yr)) { # starting at i = 2 since proj_year[1,] is given already
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,] # must compensate i
}


# The rest is wrangling & visualization (run to check it out):

# update colnames
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

# convert to df, pivot longer for 'easy rows'
proj_df <- data.frame(time_yr, proj_year) %>% 
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")

ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens! 
